An example cleaning script

### 2024-06-13 code/processing-code/processingfile.qmd
Finished most of the date cleaning and documentation.

```{r}
#install.packages(pacman)
pacman::p_load(tidyverse,lubridate,skimr,here)
```

@@

```{r}
# path to data
# note the use of the here() package and not absolute paths
data_location <- here::here("data","raw-data","raw_crime_060424.csv")
rawdata <- read.csv(data_location, TRUE)
```

@@

By inspecting the data as done above, we find some problems that need addressing:

```{r}
d1 <- rawdata %>% dplyr::filter( Height != "sixty" ) %>% 
                  dplyr::mutate(Height = as.numeric(Height))
skimr::skim(d1)
hist(d1$Height)
```

@@

>>>>

```{r}
#install.packages(pacman)
pacman::p_load(tidyverse,lubridate,skimr,here, stringr)
```

@@

```{r}
# path to data
# note the use of the here() package and not absolute paths
data_location <- here::here("data","raw-data","raw-crime-060424.csv")
rawdata <- read.csv(data_location, TRUE)
```

@@

By inspecting the data as done above, we find some problems that need addressing:\

### Dates and Times

We have six time related fields in this dataset: Occurred.Date.Time, Occurred.Date, Occurred.Time, Report.Date.Time, Report.Date, and Report.Time.\
- The Date-Time and Date variables were read in as string, and the two Time variables were read in as numeric.
So that we can treat these variables as actual representations of a point in time, we will need to re-code them into date, time, and date-time values.\
- Many of the variables have missing or empty values where they shouldn't.
For example, there are some instances where the Occurred.Date.Time variable is empty, but the Occurred.Date and Occurred.Time values are not.\
- Finally, there are some instances where the time value does not make any sense. Values like 60 or 70 are impossible to occur and somewhat vague, so special consideration should be taken to account for these as well.

```{r}
d1 <- rawdata %>% mutate(
  test = (str_sub(as.character(Occurred.Time), -2, -1)
          %>% as.numeric()),
  Occurred.Date = mdy(Occurred.Date)
  ,Report.Date = mdy(Report.Date)
  ,Occurred.Date.Time = parse_date_time(Occurred.Date.Time,'%m/%d/%Y %I:%M:%S %p')
  ,Report.Date.Time = parse_date_time(Report.Date.Time,'%m/%d/%Y %I:%M:%S %p')
  ) %>% mutate(
    Occurred.Time = if_else(
      is.na(Occurred.Time) | test >= 60, 
      if_else(
        Occurred.Date == Report.Date,
        if_else(
          is.na(Report.Time), 
          2359, Report.Time
          ), 2359
        ), if_else(
          Occurred.Time == 2400, 
          2359, Occurred.Time
          ) 
      ) %>% as.character()
    ) %>% mutate(
      Occurred.Time = paste0('000',Occurred.Time,'00') %>% 
        str_sub(.,-6,-1) %>% 
        parse_time('%H%M%S')
      ) %>% select(-test)
```

@@

```{r Sandbox}
#rawdata %>% filter(Occurred.Date.Time == '')

d1 <- rawdata %>% mutate(
  test = (str_sub(as.character(Occurred.Time), -2, -1)
          %>% as.numeric()),
  Occurred.Date = mdy(Occurred.Date)
  ,Report.Date = mdy(Report.Date)
  ,Occurred.Date.Time = parse_date_time(Occurred.Date.Time,'%m/%d/%Y %I:%M:%S %p')
  ,Report.Date.Time = parse_date_time(Report.Date.Time,'%m/%d/%Y %I:%M:%S %p')
  ) %>% mutate(
    Occurred.Time = if_else(
      is.na(Occurred.Time) | test >= 60, 
      if_else(
        Occurred.Date == Report.Date,
        if_else(
          is.na(Report.Time), 
          2359, Report.Time
          ), 2359
        ), if_else(
          Occurred.Time == 2400, 
          2359, Occurred.Time
          ) 
      ) %>% as.character()
    ) %>% mutate(
      Occurred.Time = paste0('000',Occurred.Time,'00') %>% 
        str_sub(.,-6,-1) %>% 
        parse_time('%H%M%S')
      ) %>% select(-test)
```
